---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```



<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/onetime)](https://CRAN.R-project.org/package=onetime)
[![R-universe](https://hughjonesd.r-universe.dev/badges/onetime)](https://hughjonesd.r-universe.dev/ui#package:onetime)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/hughjonesd/onetime/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/hughjonesd/onetime/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/hughjonesd/onetime/branch/master/graph/badge.svg)](https://app.codecov.io/gh/hughjonesd/onetime?branch=master)
<!-- badges: end -->

# onetime

> Listen very carefully. I shall say this only once.

[Michelle Dubois, *'Allo 'Allo*](https://www.youtube.com/watch?v=M-_5JJmNB6E)

The onetime package provides convenience functions to run R code only once 
(ever) per user. For example, you can:

* Show a startup message only the first time (ever) that a package is loaded. 
* Run cleanup code just once after an upgrade.
* Show the user a message, with the option not to show it again.

Onetime is a lightweight package. It requires just two package dependencies,
rappdirs and filelock, with no further indirect dependencies. The total 
size including these dependencies is less than 50 Kb.

* For more information, see the [vignette](https://hughjonesd.github.io/onetime/dev/articles/onetime.html).

* Development documentation is at <https://hughjonesd.github.io/onetime/dev>.  

```{r, child = "example.Rmd"} 
```

## Installation

Onetime is now on CRAN. You can install it with:

``` r
install.packages("onetime")
```

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("hughjonesd/onetime")
```

